---
title: "Project Step 3"
author: "William Cuthbert, Sebastien Montgrain, Adrian Antonelli"
format: html
---

**Project 1: Step 3**

We wanted to round out our dataset by adding information about vaccinations and mortality for Measles by region.

For our vaccination data, we utilized the WHO's data on vaccine reports from 2012-2024

```{r}
eval = FALSE

library(readxl)
library(dplyr)

#Vaccine <- read_xlsx("Measles vaccination coverage 2026-15-04 12-34 UTC.xlsx")

#Vaccine <- Vaccine |> filter(GROUP == "WHO_REGIONS") |> select("CODE", "YEAR", "ANTIGEN", "TARGET_NUMBER", "DOSES", "COVERAGE")

#write.csv(Vaccine, "Vaccine.csv")
getwd()

```

We have successfully created a csv file of the vaccination data.

Up next, we will use web scraping

```{r}
eval = FALSE

library(rvest)
library(janitor)
library(tidyr)
library(stringr)

url <- "https://www.cdc.gov/mmwr/volumes/71/wr/mm7147a1.htm"

page <- read_html(url)

tables <- page |> html_elements("table")

table1 <- tables[[1]] |> 
  html_table(fill = TRUE) |>
  clean_names()

```

Now we have our table, but it's a huge mess, as we have issues such as multiple variables being placed in the same column, and rows indicating region that span all columns.

To work through these issues, we consulted Chat GPT about how we could separate conjoined variables, and tidy up the region name rows.

Initially we cleaned up variable names

```{r}
eval = FALSE

table1 <- table1 |>
  rename(
    year_region = yr_who_region_no_of_countries_in_region,
    measles_cases_pct = percentage,
    measles_deaths_pct = percentage_2,
    mortality_reduction_pct = percentage_3,
    deaths_averted = percentage_4,
    reported_cases_share = no_of_reported_measles_cases_percent_of_total_cases,
    incidence = measles_incidence
  )

```

We then were told that "\^\\d{4}" indicates the string to detect exactly 4 digits at the beginning of the string. This is essentially capturing the year.

```{r}
eval = FALSE

table1 <- table1 |>
  mutate(
    is_year = str_detect(year_region, "^\\d{4}"),
    region = if_else(!is_year, year_region, NA_character_)
  ) |>
  fill(region)

table1 <- table1 |>
  filter(is_year)

```

The regex section of code was explained to be the line separating the year from the number of countries per region (which were initially stuck in the same column).

```{r}
eval = FALSE

table1 <- table1 |>
  extract(
    year_region,
    into = c("year", "n_countries"),
    regex = "(\\d{4}) \\((\\d+)\\)", 
    convert = TRUE
  )

table1 <- table1 |>
  mutate(across(
    where(is.character),
    ~ str_replace_all(., ",", "")
  ))

table1 <- table1 |>
  mutate(across(
    c(measles_cases_pct, measles_deaths_pct,
      mortality_reduction_pct, deaths_averted,
      reported_cases_share, incidence),
    as.numeric
  ))

```

Now we just need to filter, rename, and select relevant columns for our main data set, then save our newly configured measles coverage data.

```{r}
eval = FALSE

coverage <- table1 |> 
  filter(region != "Total (all regions)" & year != 2000) |>
  select(year, measles_cases_pct, measles_deaths_pct, mortality_reduction_pct, deaths_averted, incidence, region)


write.csv(coverage, "Coverage.csv")

```
